---
title: "Homework 3"
author: "Lara Friedlund"
date: "2/8/2019"
output: html_document
---
```{r}
library(tidyverse)
```

```{r}
fisheries <- readr::read_csv(file = "FAO_1950to2012_111914.csv")
```
# 1: It seems that there are multiple columns named species. Also, the columns with 1950-2012 can be formatted better.

# 2
```{r}
names(fisheries) = make.names(names(fisheries), unique=T)
names(fisheries)
```
```{r}
library(dplyr)
```

# 3
```{r}
fisheries<-
fisheries %>% 
  dplyr::rename(
    country = Country..Country.,
    commname = Species..ASFIS.species.,
    sciname = Species..ASFIS.species._2,
    spcode = Species..ASFIS.species._1,
    spgroup = Species..ISSCAAP.group.,
    spgroupname = Species..ISSCAAP.group._1,
    region = Fishing.area..FAO.major.fishing.area.,
    unit = Measure..Measure.)
```

# 4: The data is not tidy because variables (years) are being used as columns.

# 5
```{r}
fisheries_tidy <- 
  fisheries %>% 
  gather(num_range('X',1950:2012), key='year', value='catch')
fisheries_tidy
```

# 6: Both the year and catch columns are character class. There is missing/empty data in the catch column.
```{r}
glimpse(fisheries_tidy)
```

# 7
```{r}
fisheries_tidy <- 
  fisheries_tidy %>% 
  mutate(
    year= as.numeric(str_replace(year, 'X', '')), 
    catch= as.numeric(str_replace(catch, c(' F','...','-'), replacement = '')) 
    )
```

# 8: Data is tidy because varliables are no longer columns

# 9:

```{r}
fisheries_tidy %>% 
  select(country, spgroupname, year, catch) %>%
  filter(year>= 2008, year<=2012,
    spgroupname=="Squids, cuttlefishes, octopuses") %>%
  arrange(desc(catch))
```
# Peru, China, Vietnam, Japan, Taiwan Province of China

# 10: France
```{r}
cuttle <- 
  fisheries_tidy %>% 
  select(country, commname, year, catch) %>% 
  filter(year>=2008 & year<=2012) %>%
  filter(commname=="Common cuttlefish") %>%
  arrange(desc(catch))
cuttle
```

