---
title: "Homework 6"
author: "Lara Friedlund"
date: "March 22, 2019"
output: 
  html_document: 
    fig_caption: yes
    keep_md: yes
---
```{r}
library(tidyverse)
library(skimr)
library("RColorBrewer")
```

##Gapminder

```{r}
library(gapminder)
```

```{r}
gapminder <- gapminder::gapminder
```

#1 Explore the data using the various function you have learned. Is it tidy, are there any NA's, what are its dimensions, what are the column names, etc.

```{r}
names(gapminder)
```

```{r}
gapminder %>% skimr::skim()
```

```{r}
gapminder %>%
  purrr::map_df(~ sum(is.na(.)))%>% 
  tidyr::gather(variables, num_nas) %>% 
  arrange(desc(num_nas))
```

#2 We are interested in the relationship between per capita GDP and life expectancy; i.e. does having more money help you live longer on average. Make a quick plot below to visualize this relationship.
```{r}
gapminder %>%
  ggplot(aes(x=gdpPercap, y=lifeExp))+
  geom_point()+
  labs(title = "GDP vs. Life Expectancy", x = "GDP Per Capita", y = "Life Expectancy")+
  theme(plot.title = element_text(size = rel(1.5), hjust = 0.5))
```


#3 There is extreme disparity in per capita GDP. Rescale the x axis to make this easier to interpret. How would you characterize the relationship?
```{r}
gapminder %>%
  ggplot(aes(x=gdpPercap, y=lifeExp))+
  geom_point()+
  scale_x_log10()+
  labs(title = "GDP vs. Life Expectancy", x = "GDP Per Capita", y = "Life Expectancy")+
  theme(plot.title = element_text(size = rel(1.5), hjust = 0.5))
```

#4 This should look pretty dense to you with significant overplotting. Try using a faceting approach to break this relationship down by year.
```{r}
gapminder %>%
  ggplot(aes(x=gdpPercap, y=lifeExp))+
  geom_point()+
  facet_wrap(~year)
```

#5 Simplify the comparison by comparing only 1952 and 2007. Can you come to any conclusions?
```{r}
gapminder %>%
  filter(year==1952 | year==2007) %>% 
  ggplot(aes(x=gdpPercap, y=lifeExp))+
  geom_point()+
  facet_wrap(~year)
#As time has gone by, life expectancy has gone up. GDP for the higher life expectancy increased.
```

#6 Let's stick with the 1952 and 2007 comparison but make some aesthetic adjustments. First try to color by continent and adjust the size of the points by population. Add + scale_size(range = c(0.1, 10), guide = "none") as a layer to clean things up a bit.
```{r}
gapminder %>%
  filter(year==1952 | year==2007) %>% 
  ggplot(aes(x=gdpPercap, y=lifeExp, color=continent, size=pop))+
  geom_point()+
  facet_wrap(~year)+
  scale_size(range = c(0.1, 10), guide = "none")
```


#7 Although we did not introduce them in lab, ggplot has a number of built-in themes that make things easier. I like the light theme for these data, but you can see lots of options. Apply one of these to your plot above.
```{r}
gapminder %>%
  filter(year==1952 | year==2007) %>% 
  ggplot(aes(x=gdpPercap, y=lifeExp, color=continent, size=pop))+
  geom_point()+
  facet_wrap(~year)+
  scale_size(range = c(0.1, 10), guide = "none")+
  theme_light()
```

#8 What is the population for all countries on the Asian continent in 2007? Show this as a barplot.
```{r}
gapminder %>%
  filter(year==2007, continent=="Asia") %>% 
  ggplot(aes(x=reorder(country, -pop), y=pop))+
  geom_col(fill="green")+
  coord_flip()+
  labs(title = "Population on Asian Continent (2007)", x = "Population", y = "Country")+
  theme_light()
```


#9 You should see that China's population is the largest with India a close second. Let's focus on China only. Make a plot that shows how population has changed over the years.
```{r}
gapminder %>%
  filter(country=="China") %>% 
  ggplot(aes(x=factor(year), y=pop))+
  geom_bar(stat="identity", fill="darkgreen")+
  labs(title = "Population Growth in China", x = "Year", y = "Population")
```

#10 Let's compare China and India. Make a barplot that shows population growth by year using position=dodge. Apply a custom color theme using RColorBrewer.
```{r}
gapminder %>%
  filter(country=="China" | country=="India") %>% 
  ggplot(aes(x=factor(year), y=pop, fill=country))+
  geom_bar(stat="identity", position="dodge")+
  scale_fill_brewer(palette = "Paired")+
  labs(title = "Population Growth in China and India", x = "Year", y = "Population")
```