---
title: "Homework 5"
author: "Lara Friedlund"
date: "March 22, 2019"
output: 
  html_document: 
    fig_caption: yes
    keep_md: yes
---
---
title: "HW5"
author: "Lara Friedlund"
output: 
  html_document: 
    fig_caption: yes
    keep_md: yes
---
##Mammals Life History

#1 Load the data.
```{r}
library(tidyverse)
```

```{r}
life_history <- readr::read_csv(file = "data/mammal_lifehistories_v2.csv")
```


#2 Use your preferred function to have a look. Do you notice any problems?
```{r}
library(skimr)
```

```{r}
life_history %>%
  skimr::skim()
```


#3 There are NA's. How are you going to deal with them?
```{r}
life_history <- life_history %>% 
  na_if("-999")
```

```{r}
life_history
```


#4 Where are the NA's? 
```{r}
life_history %>%
  purrr::map_df(~ sum(is.na(.)))%>% 
  tidyr::gather(variables, num_nas) %>% 
  arrange(desc(num_nas))
```


#5 Some of the variable names will be problematic. Let's rename them here as a final tidy step.
```{r}
life_history <- life_history %>%
  dplyr::rename(genus = "Genus", wean_mass = "wean mass", max_life = "max. life", litter_size = "litter size", litters_yr = "litters/year")
```

```{r}
life_history
```

##Ggplot 

```{r}
options(scipen=999)
```

#6 What is the relationship between newborn body mass and gestation? Make a scatterplot that shows this relationship.
```{r}
life_history %>% 
  ggplot(aes(x=newborn, y=gestation))+
  geom_point()+
  labs(title = "Newborn Body Mass vs. Gestation", x = "Newborn Mass", y = "Gestation")+ 
  theme(plot.title = element_text(size = rel(1.5), hjust = 0.5))
```


#7 You should notice that because of the outliers in newborn mass, we need to make some changes. We didn't talk about this in lab, but you can use scale_x_log10() as a layer to correct for this issue. This will log transform the y-axis values.
```{r}
life_history %>%
  ggplot(aes(x=newborn, y=gestation))+
  geom_point()+
  scale_x_log10()+
  labs(title = "Newborn Body Mass vs. Gestation", x = "Newborn Mass", y = "Gestation")+
  theme(plot.title = element_text(size = rel(1.5), hjust = 0.5))
```

#8 Now that you have the basic plot, color the points by taxonomic order.
```{r}
life_history %>%
  ggplot(aes(x=newborn, y=gestation, color=order))+
  geom_point()+
  scale_x_log10()+
  labs(title = "Newborn Body Mass vs. Gestation", x = "Newborn Mass", y = "Gestation")+
  theme(plot.title = element_text(size = rel(1.5), hjust = 0.5))
```


#9 Lastly, make the size of the points proportional to body mass.
```{r}
life_history %>%
  ggplot(aes(x=newborn, y=gestation, color=order, size=mass))+
  geom_point()+
  scale_x_log10()+
  labs(title = "Newborn Body Mass vs. Gestation", x = "Newborn Mass", y = "Gestation")+
  theme(plot.title = element_text(size = rel(1.5), hjust = 0.5))
```


#10 Make a plot that shows the range of lifespan by order.
```{r}
life_history %>% 
  ggplot(aes(x=order, y=max_life, fill=order))+
  geom_boxplot()+
  coord_flip()+
  scale_y_log10()+
  labs(title = "Lifespan by Taxonomic Order", x = "Taxonomic Order", y = "Lifespan", fill="Taxonomic Order")
```

